\LARGE \bf Human-Robot Interaction: Tackling the AI Challenges

### 2013-08-30 paper.tex
Completed the 'experiment' section
Since no input from LAAS, re-using 'Moving to London' scenario.
While here, corrected/fixed/simplified the 'Cleaning the table' description

Our architecture has been deployed and tested in a number of experiments on
several robotic platforms. Table~\ref{table|experiences} lists the most
significant ones, with their main focuses and reference publication for the
interested reader.

@@

\fxfatal{Most of the Experiment section needs to be written}

We assume here that the robot (and the human) has been given the joint goal
``CLEAN TABLE''. For HATP, this means putting all tapes that are
currently on the table in the trashbin. Depending on the state of the
world and agent preferences, different plans are produced.

Figure~\ref{plan-etat2} shows a plan produced to clean the table based on the initial the initial state given in  \S~\ref{sub:gtrc}.

\begin{figure*}[thpb]
  \centering
  \includegraphics[width=1.0\textwidth]{./figs/plan3.pdf} \\
  \caption {A plan produced by HATP for clean the table based on the initial state given in \S~\ref{sub:gtrc}}
  \label{plan-etat2}
\end{figure*}

Let us now take a simpler example to illustrate a full run of the
system. We have only one tape on the table and it is is reachable only
by the robot while the throw position on top of the trashbin is
reachable only by the human.

\begin{figure*}[thpb]
  \centering
    \includegraphics[width=1.0\textwidth]{./figs/manip_run.pdf} \\
    \begin{center}
    \caption {An example of a human-robot collaborative goal achievement}
    \end{center}
  \label{manip_run_fg}
\end{figure*}

%\subsection{successful execution}

Figure~\ref{manip_run_fg} illustrates the main processes occurring
during a multi-step human-robot collaborative goal achievement.  The
plan produced is quite straightforward and is shown in the third row
called ``Goal and Plan''. It consists in 4 successive actions
involving the robot and the human. Robot grasps the tape and then
places it on the table at a position where it is visible and reachable
for the human. Human then is asked to pick the tape and throw it in
the trashbin. The first row, named ``Cameras'', shows several
snapshots corresponding to various execution steps. Snapshot~1
presents the initial situation. Snapshots~2, 3, 4 and 5 give the state
after the successive achievement of the four actions in the plan. The
second row, named ``3D Model'', shows the display of SPARK at the same
instants. The fourth, row called ``Robot Speech Acts'', gives robot
speech acts produced along the execution to inform the human partner
about goal and plan creation and status and to verbalize the actions
that the human is asked to execute. The fifth row illustrates robot
knowledge on itself and on the objects. The sixth row illustrates the
robot knowledge about the human state. The seventh row gives ongoing
robot action with action preconditions and effects assessment as well
as motion execution tasks. Motion trajectory typology can be found
between the 3D Model views. The eighth row gives ongoing human action
with action preconditions and effects assessment and monitoring
activity.

@@

The human presence brings new requirements for robot's abilities both
at the functional and at the deliberative levels~\cite{Klein2004}. The
topics involve motion~\cite{Kulic2007,Berg2004,Madhava2006},
navigation~\cite{Althaus2004,Sisbot2007}, manipulation~\cite{Kemp2007}
in presence of humans as well as perception of human
activities~\cite{Breazeal2001,Burger2008}. Also, when
interacting with humans, robots need to incorporate communication and
collaboration abilities. Several theories dealing with
collaboration~\cite{Cohen1991,Grosz1996,Clark1996} emphasize that
collaborative tasks have specific requirements compared to individual
ones, \eg, since the robot and the person share a common goal, they
have to agree on the manner to realize it, they must show their
commitment to the goal during execution, etc. Several robotic systems
have already been built based on these
theories~\cite{Rich1997,Sidner2005,Tambe2005a,Breazeal2003} and they
all have shown benefits of this approach\fxwarning{which approach? how it is related to larger AI challenges?}. They have also shown how
difficult it is to manage turn-taking between communication partners
and to interleave task realization and communication in a generic
way\fxwarning{is it smthg we improve one?}. Finally, today only few
systems~\cite{Fong2006,Breazeal2003,Sisbot2008} take humans into
account at all levels\fxwarning{Old references, no new ones? we should also say what is new/better with our approach (Rachid)}.

>>>>

Our architecture has been deployed and tested in a number of experiments on
several robotic platforms. Table~\ref{table|experiences} lists the most
significant ones, with their main focuses and reference publications for the
interested reader.

@@

The two following sections present one such experiment (the {\it Moving to
London} scenario), and the synopsis of another one (the {\it Cleaning the
table} scenario).

The first one is focused on knowledge representation and verbal interaction:
the human asks for help to find and pack objects. In this experiment, robot
actions are limited to simple motions (like head tracking or predefined
pick-and-place).

The second one involves the SHARY execution controller and the HATP symbolic
task planner besides ORO and {\sc Dialogs}. In this scenario, the human and the
robot try to cooperatively remove objects from a table. In this scenario,
behaviours and motions are fully planned.

\subsection{\emph{Moving to London} scenario}

This first experiment is based on the following daily life situation: Tom and
Jerry are moving to London, so they are packing things in boxes. The scenario
takes places in the living-room, where Jido (our robot) is observing while they
move things here and there. To assess the reasoning abilities of the robot they
ask Jido for information (entered through keyboard). Ideally, the robot should
also perform actions when required (e.g. hand an object when asking ``give
me...''). However, since it is out of the scope of this work, we do not include
any motion from the robot's side.

Perception of objects is done through a tag-based system and humans are
detected through motion capture. The robot knowledge base is pre-loaded with
the \emph{ORO Commonsense Ontology}.  We next describe in detail two
situations where we can follow the internal robot's reasoning and the
interaction with the users.

\paragraph{Implicit disambiguation through visual perspective taking}

\begin{figure}[!ht]
  \centering
  \includegraphics[width=0.9\linewidth]{./figs/pt.jpg}
\caption{Asking the robot to hand over a tape on the table.}
  \label{fig|vpt}
\end{figure}

Tom enters the room while carrying a big box (Figure~\ref{fig|vpt}, page 1). He
approaches the table and asks Jido to handle him the video tape: ``Jido, can
you give me the video tape''. The \textsc{Dialogs} module queries the ontology to
identify the object the human is referring to: \stmt{?obj type VideoTape}.

There are two video tapes in the scene: one on the table, and another one
inside the cardboard box. Thus, the knowledge base returns both: $\Rightarrow$
\concept{?obj = [videoTape1, videoTape2]}.

However, only one is visible for Tom (the one on the
table). Thus, although there is an ambiguity from the robot's perspective
(since it can see both video tapes), based on the perspective of its human
partner it infers that Tom is referring to the video tape on the table, and not
the one inside the box which is not visible from his view. Therefore,
non-visible objects are removed obtaining: \concept{?obj =[videoTape1]}.

Since only one object is available, the robot infers
that the human refers to it and would eventually execute the command, \ie give
it to the human. Alternatively, the robot could first verify with the human if
that was the object being referred to or not before proceeding to execute the
action. Table~\ref{table|ptbeliefs} lists the robot's beliefs about itself and
its human partner involved in this situation.

\begin{table}
\begin{center}
\begin{tabular}{l}
\hline
Robot's beliefs about itself (\emph{robot's model}):\\
\hline
  \hspace{0.7cm}\stmt{videoTape1 type VideoTape}\\
  \hspace{0.7cm}\stmt{videoTape1 isOn table}\\
  \hspace{0.7cm}\stmt{videoTape1 isVisible true}\\
  \hspace{0.7cm}\stmt{videoTape2 type VideoTape}\\
  \hspace{0.7cm}\stmt{videoTape2 isIn cardBoardBox}\\
  \hspace{0.7cm}\stmt{videoTape2 isVisible true}\\
\hline
\hline
Robot's beliefs about Tom (\emph{Tom's model}):\\
\hline
  \hspace{0.7cm}\stmt{videoTape1 type VideoTape}\\
  \hspace{0.7cm}\stmt{videoTape1 isOn table}\\
  \hspace{0.7cm}\stmt{videoTape1 isVisible true}\\
  \hspace{0.7cm}\stmt{videoTape2 type VideoTape}\\
  \hspace{0.7cm}\stmt{videoTape2 isIn cardBoardBox}\\
  \hspace{0.7cm}\stmt{videoTape2 isVisible false}\\
 \hline
\end{tabular}
\end{center}
\caption{Robot's beliefs about itself and its human partner.}
\label{table|ptbeliefs}
\end{table}

\paragraph{Explicit disambiguation through verbal interaction and gestures}

\begin{figure}[!ht]
  \centering
  \includegraphics[width=0.9\linewidth]{./figs/inTheBox2.jpg}
\caption{Jerry asks Jido for the content of the box by pointing at it.}
  \label{fig|pointing}
\end{figure}

In this situation, Jerry enters the
living room without knowing where Tom had placed the video tapes. So he first
asks Jido: ``What's in the box?''. Before the robot can answer the question it
has to figure out which box Jerry is talking about. Similar to the previous
situation, there are two available boxes:

\begin{center}
\begin{tabular}{l}
\stmt{?obj type box}\\
\hspace{0.7cm}$\Rightarrow$ {\tt ?obj = [cardBoardBox, toolbox]}
\end{tabular}
\end{center}

However both are visible and the cognitive ambiguity resolution cannot be
applied. The only option is to ask Jerry which box he is referring to: ``Which
box, the toolbox or the cardboard box?'' Jerry could now simply answer the
question. Instead, he decides to point at it while indicating: ``This box''
(Figure~\ref{fig|pointing}). The robot's perception identifies the {\tt
cardBoardBox} as being pointed at and looked at by the human and updates the
ontology with this new information using a rule available in the commonsense
ontology ({\tt \textbf{pointsAt}(?ag, ?obj) $\land$ \textbf{looksAt}(?ag, ?obj) $\to$
\textbf{focusesOn}(?ag, ?obj)}) The \textsc{Dialogs} module is then able to merge both
sources of information, verbal (``this'') and gestural to distinguish the box
Jerry refers to.

\begin{center}
\begin{tabular}{l}
\stmt{Jerry pointsAt carboardBox}\\
\stmt{Jerry looksAt carboardBox}\\
$\to$ \stmt{Jerry focusesAt carboardBox}\\
\hspace{0.7cm}$\Rightarrow$ {\tt ?obj = [cardBoardBox]}
\end{tabular}
\end{center}

Finally, the \textsc{Dialogs} queries the ontology about the content of the box
and the question can be answered: ``Jido-E''. Note that the object's label is
used instead of its ID. This way we enhance interaction using familiar names
given by the users.

\begin{center}
\begin{tabular}{l}
\stmt{?obj isIn cardBoardBox}\\
\hspace{0.7cm}$\Rightarrow$ \concept{?obj = videoTape2}\\
\end{tabular}
\end{center}

At this point Jerry wants to know where the other tape is, and that is exactly
what he asks Jido: ``And where is the other tape?''. In this occasion, the
\textsc{Dialogs} module is able to interpret that Jerry is not referring to the
video which they were just talking about, but to the other one:

\begin{center}
\begin{tabular}{l}
\stmt{?obj type VideoTape}\\
\stmt{?obj differentFrom videoTape2}\\
\hspace{0.7cm}$\Rightarrow$ \concept{?obj = [videoTape1]}
\end{tabular}
\end{center}

Since there is only one possible ``other'' video (there are only two videos in
the scene), it can directly answer Jerry: ``The other tape is on the table and
next to the toolbox.''

\begin{center}
\begin{tabular}{l}
\stmt{videoTape1 isOn table}\\
\stmt{videoTape1 isNextTo toolbox}
\end{tabular}
\end{center}

\subsection{\emph{Cleaning the table} scenario}

This experiment involved a more complex decisional layer where the ORO
server was used in conjunction with the HATP symbolic task planner and the
SHARY execution controller (the complete software architecture we deployed for
this experiment is pictured on Figure~\ref{fig|archi},
page~\pageref{fig|archi}).

In this scenario (Figure~\ref{fig|cleantable-video}), a human and a robot
cooperate to remove objects from a table. The robot produces symbolic plans for
both itself and the human (Figure~\ref{plan-etat2})
that allow the robot to (verbally) share the task with the human (like ``I take
the green box and I put it in the trashbin, you take the black video tape and
you throw it in the trashbin''). Plans are created based on perceived
visibility and reachability of the objects, and the robot also monitors the
human activities to track the advancement of the whole plan (this last aspect
is presented in~\cite{warnier2012when}).

\begin{figure}
    \centering
    \includegraphics[width=0.7\columnwidth]{./figs/cleantable.jpg}

    \caption{Snapshot of the film of the ``Clean the Table'' scenario. The
    physical situation, the SPARK model, and the current step of the plan can
    be seen.}

    \label{fig|cleantable-video}
\end{figure}

\begin{figure*}[thpb]
  \centering
  \includegraphics[width=1.0\textwidth]{./figs/plan3.pdf} \\
  \caption {A plan produced by HATP to execute the high-level order ``clean the table''. Two streams of actions are generated: for the human (top) and for the robot(bottom). Synchronization points ensure the coordination.}
  \label{plan-etat2}
\end{figure*}

Figure~\ref{fig|cleantable-timeline} presents a excerpt of the whole task and
shows how the different components produce and use symbolic knowledge (note
that the plan is actually computed {\it a priori}. It appears nevertheless on
this diagram for better readability).

This sequence of events depicts a run with only one tape on the table. The tape is
reachable by the robot only, while the human is the one who can reach the
trashbin to throw objects.

\begin{figure*}[thpb]
  \centering
    \includegraphics[width=1.0\textwidth]{./figs/manip_run.pdf} \\
    \begin{center}
        \caption {Sequence of events of a simplified version of the \emph{Cleaning the table} scenario.}
    \end{center}
  \label{fig|cleantable-timeline}
\end{figure*}

The plan produced in that case by HATP is straightforward and shown in the
third row. It consists in four successive actions involving the robot and the
human. Robot grasps the tape and then places it on the table at a position
where it is visible and reachable for the human. Human then is asked to pick
the tape and throw it in the trashbin.

@@

The human presence brings new requirements for robot's abilities both
at the functional and at the deliberative levels~\cite{Klein2004}. The
topics involve motion~\cite{Kulic2007,Berg2004,Madhava2006},
navigation~\cite{Althaus2004,Sisbot2007}, manipulation~\cite{Kemp2007}
in presence of humans as well as perception of human
activities~\cite{Breazeal2001,Burger2008}. Also, when
interacting with humans, robots need to incorporate communication and
collaboration abilities. Several theories dealing with
collaboration~\cite{Cohen1991,Grosz1996,Clark1996} emphasize that
collaborative tasks have specific requirements compared to individual
ones, \eg, since the robot and the person share a common goal, they
have to agree on the manner to realize it, they must show their
commitment to the goal during execution, etc. Several robotic systems
have already been built based on these
theories~\cite{Rich1997,Sidner2005,Breazeal2003} and they
all have shown benefits of this approach\fxwarning{which approach? how it is related to larger AI challenges?}. They have also shown how
difficult it is to manage turn-taking between communication partners
and to interleave task realization and communication in a generic
way\fxwarning{is it smthg we improve one?}. Finally, today only few
systems~\cite{Fong2006,Breazeal2003,Sisbot2008} take humans into
account at all levels\fxwarning{Old references, no new ones? we should also say what is new/better with our approach (Rachid)}.
